Project 1
=========


1:Loading the data into R
```{r}
setwd("C:/Users/peter/Desktop/R/Data/Reproducible Research")
activity<- read.csv("activity.csv")
```


2:Histogram of total number of steps taken each day
===
```{r}
dailysteps<-aggregate(steps~date, activity, sum, na.rm=T)
    hist(dailysteps$steps,
         breaks = 25, 
         main = "Daily Steps",
         xlab = "Number of Steps",
         col = "blue")
```



3:Mean and median number of steps taken each day
===
```{r}
mean<-mean(dailysteps$steps)
median<-median(dailysteps$steps)
```
The mean number of steps taken each day is `r mean`  
The median number of steps taken each day is `r median`



4:Time series plot of the average number of steps taken
===
```{r}
StepsPerInterval <- tapply(activity$steps, 
                           activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(StepsPerInterval)), 
     StepsPerInterval, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Daily Activity Pattern", 
     type = "l",
     col="blue")
```




5:The 5-minute interval that, on average, contains the maximum number of steps
===
```{r}
AvgMaxInter<-names(sort(StepsPerInterval, decreasing = TRUE)[1])
MaxSteps <- sort(StepsPerInterval, decreasing = TRUE)[1]
```
The intervall containing, on average, the maximum number of steps is `r AvgMaxInter`.  
It averages `r MaxSteps` steps
  
    
6:Code to describe and show a strategy for imputing missing data
===
First we assess the number of NA's in the dataset
```{r}
 nas<-sum(is.na(activity$steps))
```
There are `r nas` missing values in this dataset
These values for the intervals with missing values will be replaced by  
the average number of steps per interval. The next step will compute this average and create a new dataset with the average replacing NA's.
```{r}
StepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
activity.split <- split(activity, activity$interval)
for(i in 1:length(activity.split)){
    activity.split[[i]]$steps[is.na(activity.split[[i]]$steps)] <- StepsPerInterval[i]
}
activity.imputed <- do.call("rbind", activity.split)
activity.imputed <- activity.imputed[order(activity.imputed$date) ,]
```
Create a histogram of the total number of steps taken each day,according to the new dataset
```{r}
StepsPerDay.imputed <- tapply(activity.imputed$steps, 
                              activity.imputed$date, sum)
hist(StepsPerDay.imputed, 
     xlab = "Number of Steps", 
     main = "Histogram: Steps per Day (Imputed data)")
```
  
Calculate the mean and median steps per day of the Imputed data

```{r}
imp.daily.mean <- mean(StepsPerDay.imputed, na.rm = TRUE)
imp.daily.median <- median(StepsPerDay.imputed, na.rm = TRUE)
```
The mean number of steps in the imputed dataset is `r imp.daily.mean`.
  
The median number of steps in the imputed dataset is `r imp.daily.median`



7:Are there differences in activity patterns between weekdays and weekends?
===
Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activity.imputed$day <- ifelse(weekdays(as.Date(activity.imputed$date)) 
                               == "Saturday" | weekdays(as.Date(activity.imputed$date)) 
                               == "Sunday", "weekend", "weekday")
```
Make a panel plot containing a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
StepsPerInterval.weekend <- tapply(activity.imputed[activity.imputed$day == "weekend" ,]$steps, activity.imputed[activity.imputed$day == "weekend" ,]$interval, mean, na.rm = TRUE)

StepsPerInterval.weekday <- tapply(activity.imputed[activity.imputed$day == "weekday" ,]$steps, activity.imputed[activity.imputed$day == "weekday" ,]$interval, mean, na.rm = TRUE)
par(mfrow=c(1,2))
plot(as.numeric(names(StepsPerInterval.weekday)), 
     StepsPerInterval.weekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekdays)", 
     type = "l")

```
